Browsernative Microservices

### 2017-03-01 build/paper.tex
pdf/tex build

\title{Browsernative Microservices}
\providecommand{\subtitle}[1]{}
\subtitle{Modular web architecture through new W3C specifications}
\author{Jan Peteler, FH Würzburg-Schweinfurt, jan.peteler@student.fhws.de}
\date{Januar 2017}

\begin{document}
\maketitle
\begin{abstract}
Building complex web applications nowadays require additional layers of
abstraction and often heavily depend on indispensable frameworks. While
they perfectly circumstance the global paradigm of the DOM they remain
highly proprietary attached to the framework. New w3c specifications
build right into the browser provide a native service API in the need to
create standardized web components for the browser. As of 2017 all major
browsers will ship those technologies in their browsers. The paper not
only provides an introduction to the specification. As an assessment
criteria of it will match those specifications against the concept of
microservices which is an modular system architecture to build scalable
applications.
\end{abstract}

@@

Computers can scale, humans can't. Ever since a complex system made by
humans has been constrained by humans mental capabilities. Like in the
analogy of juggling balls our brain can just ``juggle'' a few things at
a time. Rich Hickey, the inventor of the programming language Clojure
gave an inspirational keynote on the topic of
\textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
  Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} In every sphere
of a humans life simplicity aligns perception with our mental
capacities.

Derived from the Latin word \textbf{simplex} simple can be understood as
``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
  Dictionary}} which points towards an unidimensional state. While
complexity describes the multilayered und entangled nature of conditions
simplicity empowers the human brain to reason about issues
straightforward. It certainly has some overlapping's with easy, but
while easy is more of a relative spirit, simple can be laid out as a
objective manner and therefore universally applicable.

Software development is undoubtedly rich in complexity and full of
subtle pitfalls. In a typical scenario a growing software project
evolves in one or another opinionated direction over time. Layers of
abstractions wrestling with aged legacy code requiring layers of
middleware. Mutating assets create subtle bugs and so forth. Eventually
the small piece of software may end up in a highly complected monolith
which will determine future design decisions to a painful degree.
Opaqueness of the system will slow down innovation to a minimum in the
need of ``keeping the lights on''.

On the other side of this dystopian scenario a truly modular system
architecture abandons many of those potential inconsistencies. The whole
system is divided in pluggable parts, object mutation is either
traceable or avoided altogether in favor of immutable data structures.
As Rich Hickey argues, design decisions should be made under the
\textbf{impression of extending, substitution, moving, combining and
repurposing}. The ability to reason about the program at any given time
is crucial for future decisions and implementations. Recalling again the
unidimensional nature of simplicity.

Simplicity in ``the web'', read as a loose generalization of
``everything that runs in the browser'' is certainly a story full of
misconceptions. While simplicity in the backend is mostly a matter of
principles and patterns, any browser-based frontend is restricted on the
highly deterministic nature of the browser platform.

In the last four years the average transfer size of a webpage doubled to
currently around 2.5 MB.\footnote{\href{http://httparchive.org/trends.php}{HTTPArchive
  Trends}} Leaving images, fonts or other content aside the size of
HTML, CSS and JS sums up to an total average of 550 KB. One character
weights around 1 byte which means an average webpage is delivering
550.000 character or around 125 pages of single-spaced text. Frederic
Filloux analyzed the payload on different newspaper websites and came to
the conclusion, that only round about 5-6 \% of the transferred
characters made for human consumption.{[}1{]}

Having an 95 \% overhead is rather undesirable for both the consumers
and creators of the website. Since it's a widespread problem without a
single point of failure one can argue the platform itself is the
failure. By design, every pageload results in a monolithic DOM tree
managed by the browser engine. Whether rendering just a bunch of static
text nodes or an ever changing webapp the underlying global nature of
the DOM tree remains the same. Every additional piece of code added to
the webpage will invisibly add another fold of complexity to this global
object.

In an non-deterministic runtime environment, encapsulation and
modularization is a typical pattern to make complexity manageable and
accommodate future uncertainty.{[}2, p. 1{]} Since years the average JS
payload is steadily rising which can be interpreted as a trend towards
more dynamic websites. The demands to the browser platform changed from
a static page renderer to a \textbf{dynamic UI machine} without changing
the underlying architecture significantly. Under the current situation
only additional layers of abstraction can wrestle complexity.

In the recent years many \textbf{frameworks}, libraries and
methodologies approached the global nature of the DOM by scoping assets
and design rules into maintainable components. While the DOM can't be
scoped, JS can. Many frameworks, like React, Angular or Vue just to name
a few, ditched the old rule of separated HTML, CSS and JS in favor of an
additional layer of abstracted JS components (containing structure,
behavior and sometimes styling). Quiet often those frameworks mimic a
MVC pattern on top of the browser engine which is a reasonable simple
design pattern to build UIs. While frameworks are a valid approach for
building scalable web applications they remain highly opinionated embody
inherent complexity themselves and can change and break over time.
Another downside is code inflation which is a crucial point for
performance. All of those bottlenecks in the web demand for a new
standardized way for creating and evolving complex web services.

In the year 2013 thinkers, creators and browser vendors joined together
to propose \emph{The Extensible Web Manifesto}.\footnote{\href{https://extensiblewebmanifesto.org/}{The
  Extensible Web Manifesto}} The manifestos claim was to enhance the
current web platforms with new low-level capabilities. Those features
should empower creators of the web to write more declarative code and
abandon known problems and artificial abstractions. Four years later,
the enhancement of JavaScript leapfrogged and many new low-level APIs
brought to life. With this new APIs at hand a vivid web developer can
create robust websites with less code and less additional libraries.
This paper is an approach to unfold these \textbf{browsernative}
technologies to create overall simple and resilient
\textbf{microservices} for the browser

@@

In search of a better, simpler web architecture one might look on
already established pattern that proofed to fulfill enterprise needs.
Microservices are a good approach for tearing big monolithic systems
into fine-grained simple services with explicit defined boundaries. In a
nutshell a microservice is a small, autonomous service that works
together with other services seamlessly.{[}3, p. 2{]} Or with the words
of Fowler and Lewis: ``\ldots{} the microservice architectural style is
an approach to developing a single application as a suite of small
services, each running in its own process and communicating with
lightweight mechanisms, often an HTTP resource API.''{[}4{]} Yet at this
point a reader might spot some similarities with microservices and the
browser-based development: Both wrestling the problem of monolithic
architecture and both using lightweight communication mechanisms. In
fact, many big companies of ``the web'' like Amazon or Netflix
successfully transformed their monolithic system into a service based
system which gives a taste of the power behind microservices.{[}4{]}

Microservices incorporate a wide array of ideas from developing scalable
software, like \textbf{domain-driven design} where it pursues the
incorporation of real world structure in the code.{[}3, p. 2{]} Or
making use of \textbf{continuous delivery} for pushing software rapidly
through \textbf{automated deployment} mechanisms into production.{[}4{]}
Furthermore, microservices transcendent the technical perspective and
reaches into the team organization. As a primary source of truth this
paper relies on the work of Sam Newman, who has written a comprehensive
guide towards microservices{[}3{]} and the work of Fowler and
Lewis{[}4{]}. The purpose of this section is to gain confidence about
the microservices architecture in the context of the browser platform.

\subsection{2.1. Componentization via
Services}\label{componentization-via-services}

``A~\textbf{component}~is a unit of software that is independently
replaceable and upgradeable.'' {[}4{]} Components are the building
blocks of microservices. And microservices are the building blocks of
applications. Essentially the difference between microservices and
components is just the level of abstraction. Whether a concrete
microservice or a much more generic component, both share a similar set
of principles. Therefore this paper referring to both parts when talking
about \textbf{services}.

The first principle of services is the \textbf{loose coupling
principle}: changing and deploying one service shouldn't result in
changing other parts of the system.{[}3, p. 30{]} Picking CSS for
example, where variables tend to be shadowed by higher specific values
making it increasingly difficult to keep changes ought to only affect
one place in the application. A \emph{browsernative microservice}
therefore pushing encapsulation and avoiding variable mutations outside
its scope as much as possible. One solution would be scoped CSS / JS to
avoid variables leaking into the global namespace as well as hidden
implementation details to avoid mutating values.

The second principle of services is the \textbf{high cohesion
principle}: Whether designing a microservice or it's components we want
related behavior sit together and unrelated behavior to sit
elsewhere.{[}3, p. 30{]} High cohesion can be enhanced towards the more
dynamic \textbf{Single Responsibility Principle}: ``Gather together
those things that change for the same reason and separate those things
that change for different reasons.''{[}5{]} In a very quick and dirty
code quality analysis, the quality can be measured just by counting the
places changes in the code occur in order to implement a functionality.
An arbitrary threefold MVC system should require a maximum of three
changes to implement or change functionalities. The problem in browser
based development is not only the global paradigm which makes changes
deliberately unpredictable. The high cohesion principle is violated by
the traditional separation along the siloed entities \textbf{HTML, JS
and CSS}. Understanding the relation of HTML markup and another CSS file
adds incidental complexity. Different approaches emerged over the recent
years to join the forces. Most notable the dedicated frontend language
\emph{elm}\footnote{\href{http://elm-lang.org/}{elm lang}} where MODEL,
VIEW and CONTROLLER sit in one place. A \emph{browsernative service}
sought for an combination of the web native trinity HTML, JS and CSS.

Traditional web development relies on \textbf{libraries} to enhance the
service capabilities of the web platform. Compared to libraries a
component service offers multiple advantages for building, deployment
and shipping. As expressed earlier a \emph{components for the web}, or
web component, is \textbf{self-contained} which means it embodies all
needed functionality to get it's job done. Therefore it has a much
better evolution mechanism in the service contracts. Changing
functionality won't break other services. A component can progressively
enhanced which guarantees functionality throughout different versions. A
library is only loosely coupled to the implementation and therefore hard
to track in functionality. Changing a library may result in an
unforeseen amount of time fixing implementations. It is not unusual to
see websites embodying different versions of the same library to
guarantee functionality but lowers page performance significantly.

Another issue where web components stand out is related to performance
and especially the critical first page rendering. Libraries for the
browsers are traditionally ``shipped'' as non static immediately-invoked
functions run immediately on page load. Following the \emph{Google RAIL}
model, a user-centric performance measurement, a page load ought be less
than 1 second to catch up user's attention.{[}6{]} There are many ways
for optimizing the critical first render but as a rule of thumb a
build-in web component might be always superior to libraries in terms of
first rendering. The fine-grained lifecycle methods which will be
described in the technical section of this paper give the developer far
reaching optimization opportunities.

The last argument in favor of components over libraries is the more
explicit interface.{[}4{]} While the functionality of a library needs
documentation to be accessible a component functionality is exposed via
the components' signature. The markup language HTML is by design
equipped with an expressive syntax and steering a web component using
attributes and values should be most straightforward even to
unexperienced web developers.

\subsection{2.2. Organized around Business
Capabilities}\label{organized-around-business-capabilities}

\begin{quote}
``organizations which design systems \ldots{} are constrained to produce
designs which are copies of the communication~structures~of these
organizations''. {[}7{]}
\end{quote}

Emphasizing the human factor in microservices is a key feature.
Microservices are a product of real-world usage.{[}3, p. 1{]} Instead of
splitting team structures along the technology stack (UI Experts
-\textgreater{} Middleware -\textgreater{} Database) a microservice
approach model teams around \textbf{business capabilities}.{[}4{]}
Consequently every team is capable of planning, designing, implementing,
testing and maintaining their very own microservice. Along the
technology stack every member gains high competence about the the
service architecture which can be critical for evolving the service over
its whole lifecycle.

Real-world domains tend to be complex and multifaceted. To unfold their
complexity, domains can subdivided into \textbf{bounded context}.{[}3,
p. 31{]} For example, customer service is a business domain but with
varying bounded context. One context can be sales, another context could
be support. Every context makes different assumptions about the
underlying model and draws an explicit interface where it decides what
to share with other contexts.{[}3, p. 30{]} Evaluating each bounded
context within each business domain will eventually shape the data
persistency model likewise the interface to the service. This
methodology can be iterated over and over again. A sales service for
example might be evaluated to different sales context resulting in
differing interfaces on differing devices shaped by browsernative
microservices. By focusing the service into clear defined business
boundaries, it is easier to define a smart API of the service.

Assigning service responsibility to a team, the so called
\textbf{Definition of Done} (DoD) shifts from ``accomplishing projects''
to ``accomplishing products''. This new paradigm not only changes the
administrative overhead like budgeting or resource allocation. It
creates a kind of responsibility connection from the team to the service
which can be best described as \textbf{shared governance} model. ``Each
team collectively share responsibility for evolving the technical vision
of the system.''{[}3, p. 247{]} Expectedly, those teams are more
motivated within their very own service and exhibit a more sophisticated
iteration time.{[}4{]}

For many companies working in the spheres of the internet the client
side is highly important for their business. In fact, business goals and
capabilities can be derived from frontend needs. The state of the web is
not only a story of numerous artifacts, it is also a story of an highly
fragmented market along devices, operating systems, differing sizes and
functionalities. Different devices again have different assumptions
about the technology stack. Splitting teams along the stack results in
an slow paced back and forth negotiation for every change to be made
make iteration time expensive. As browser technologies, design
guidelines and devices change frequently it makes absolutely sense to
shift responsibility towards the teams altogether.

@@

To ensure the microservice functionality among teams and different
services requires thoughtful decentralization. Emphasizing once more the
real-world capabilities of microservices a message channel architecture
can be derived from patterns known from traditional postal services. A
physical letter has only two smart endpoints entitled to read and
process the message while packaging likewise the connection itself is
maximized towards unification. \emph{Smart endpoints and dumb pipes} is
coined to the approach of designing communication mostly decoupled and
as cohesive as possible.{[}4{]} Applying this rules to the web platform
can result into building unified JSON message objects passed along
``dumb'' middleware components. A typical browser event comes close to
this definition and suits arguably well for in-memory communication
between web components and microservices.

Microservices heavily rely on simple HTTP request-response with resource
APIs and lightweight messaging.{[}4{]} Newman puts his recommendation on
technologic-agnostic REST APIs to free data persistence from
implementation constraints.{[}3, p. 247{]} The advantage of this overall
simple communication model is the suitability for both frontend-backend
likewise backend-backend communication. A service therefore can evolve
from an heavy backend with a lot of network roundtrips to a leaner
backend seamlessly. The browser build-in \textbf{fetch API} which is
essentially a HTTP request can be heavily incorporated into a
browsernative microservice to ensure communication to services in the
backend.

\subsection{2.4. Decentralized
Governance}\label{decentralized-governance}

Microservices are separate entities and decentralization is important to
ensure autonomy. This paper already described fragmented services
bounded to singular business context choreographed by simple
communication protocols developed and evolved by autonomous teams.

This distributed nature empowers teams to create their own technology
stack, tools and services designed in the spirit of language- and
platform independence and share their knowledge with other
parties.{[}4{]} In the recent years many big companies like Facebook,
Google, Netflix and others followed that spirit and published their
ideas and implementations open source. The previously mentioned ReactJS
for example is a product of Facebooks need to ensure a consistent
frontend experience. In fact, many tools and techniques are byproduct of
vital interaction of concrete domain problems and their implementations.

The spirit of freedom can't be applied universally to
\emph{browsernative microservices} as the browser and its underlying DOM
will be the limitation factor to a certain degree. Talking about the
browser, a reader might be tempted to narrowly thinking of the obvious
VIEW layer only - which is not true anymore. In the recent years the
major browser engines grow to to a fully-fledged app deployment platform
offering connectors to build-in databases, multithreading support and
ever-growing JS build-ins like speech synthesis or push notifications.
So-called \textbf{Progressive Web Apps}\footnote{\href{https://developers.google.com/web/progressive-web-apps}{Progressive
  Web Apps}}, a bunch of criteria for building good browser apps, can
achieve a similar look and feel like native apps. And last but not least
services like NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}}
effectively compiling ``the web'' to native machine code lowering the
boundary between native and browser code even further.

JS is the widely accepted language of the web. Nevertheless, a
microservice engineering team might choose another language for various
reasons. Transpiling languages to JS as target language is a stable
solution nowadays. Languages like TypeScript, ClojureScript or
PureScript compile to JS even exclusively. Once web components hit a
critical mass there will be most likely some library support or foreign
function interface towards ES6 modules (which are mandatory for the new
specifications).

Another more real life decentralization aspect derives from the easiness
of deployment in a safe, sandboxed environment. Web components virtually
ship no overhead or require dedicated build tools. This makes them ideal
candidates for sharing and open source publishing similar to the largest
JS package registry NPM. In the spirit of NPM web components can be
perceived as frontend packages with an HTML interface instead of a JS
signature. Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
is a registry for ready-to-use components of every scale and purpose
where for even Google shares a lot of their material design elements.

\subsection{2.5. Decentralized Data
Management}\label{decentralized-data-management}

Data Management in a microservice follows the same modular philosophy
like the service implementation. As mentioned earlier different bounded
contexts make different assumptions of the underlying models. A
\emph{browsernative microservices} takes this idea even further and
expands it to the fragmented world of electronic devices. Decentralized
decisions about conceptual models demand for decentralized data storage
decisions.{[}4{]} Todays web architectures aims to leverage an
increasing amout of processing to the client to avoid time-consuming
roundtrips especially in mobiles networks.\footnote{Latency numbers:
  https://gist.github.com/jboner/2841832} Since network roundtrips are
costly it is a good advice to only query as much data as needed and
cache as much as possible. The build-in LocalStorage or its successor
IndexedDB are mature persistence technologies and libraries like
PouchDB\footnote{\href{https://pouchdb.com/}{PouchDB}} even offer
backend adapters for syncing out of the box.

``Microservices prefer letting each service manage its own
database.''{[}Fowler2014{]} Ben Issa, chief architect of ING Australia
emphasizes this pragmatism on APIs in a conference talk. At ING the
frontend demands tailor the backend APIs, APIs may be produced
automatically and not even Issa knows how many APIs exists.{[}8{]} They
are using a pattern called \textbf{backend for frontends} empowering the
team working to craft their UI and backend in a one-to-one
relationship.{[}3, p. 72{]}

>>>>

\title{Browsernative Microservices}
\providecommand{\subtitle}[1]{}
\subtitle{Modular web architecture through new W3C specifications}
\author{Jan Peteler, FH Würzburg-Schweinfurt, jan.peteler@student.fhws.de}
\date{March 2017}

\begin{document}
\maketitle
\begin{abstract}
Building complex web applications nowadays require additional layers of
abstraction and often heavily depend on indispensable frameworks. While
they elegantly circuit the global paradigm of the DOM they remain highly
proprietary pieces of technology. New w3c specifications build right
into the browser provide a native service API in the need to create
standardized web components for the browser. As of 2017 all major
browsers will ship those technologies in their browsers. The paper not
only provides an brief and practical introduction to the specification.
As an assessment criteria this paper will match those specifications
against the concept of microservices, which is an modular system
architecture to build scalable applications.
\end{abstract}

@@

Computers can scale, humans can't. Ever since complex systems made by
humans have been constrained by humans mental capabilities. Like in the
analogy of juggling balls our brain can just ``juggle'' a few items at a
time. Rich Hickey, the inventor of the programming language Clojure gave
an inspirational keynote on the topic of \textbf{simplicity}.{[}1{]} In
every sphere of a human's life simplicity aligns perception with mental
capacities.

Derived from the Latin word \textbf{simplex}, \emph{simple} can be
understood as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
  Dictionary}} which points towards an unidimensional state. While
complexity describes the multilayered und entangled nature of conditions
simplicity empowers the human brain to reason about issues
straightforward. It certainly has some overlappings with \emph{easy},
but while \emph{easy} is more of a relative spirit, \emph{simple} can be
laid out as an objective manner and therefore universally applicable.

Software development is undoubtedly rich in complexity and full of
subtle pitfalls. In a typical scenario a growing software project
evolves in one or another opinionated direction over time. Layers of
abstractions wrestle against aged legacy code requiring additional
middleware. Mutating assets create subtle bugs and so forth. Eventually
the small piece of software may end up in a highly complected monolith
which will determine future design decisions to a painful degree.
Opaqueness of the system will slow down innovation to a minimum in the
need of `keeping the lights on'.

On the other side of this dystopian scenario a truly modular system
architecture abandons many of those potential inconsistencies. The whole
system is divided in pluggable parts, object mutation is either
traceable or avoided altogether in favor of immutable data structures.
As Rich Hickey argues, design decisions should be made under the
\textbf{impression of extending, substitution, moving, combining and
repurposing}.{[}1{]} The ability to reason about the program at any
given time is crucial for future decisions and implementations,
recalling the unidimensional nature of simplicity.

Simplicity in `the web', read as a loose generalization of `everything
that runs in the browser' is certainly a story full of misconceptions.
While simplicity in the backend is mostly a matter of principles and
patterns, any browser-based frontend is restricted on the highly
deterministic nature of the browser platform.

In the last four years the average transfer size of a webpage doubled to
currently around 2.5 MB.\footnote{\href{http://httparchive.org/trends.php}{HTTPArchive
  Trends}} Leaving images, fonts or other content aside, the size of
HTML, CSS and JS sums up to a total average of 550 KB. One character
weights around 1 byte which means an average webpage is delivering
550.000 characters or around 125 pages of single-spaced text. Frederic
Filloux analyzed the payload on different newspaper websites and came to
the conclusion, that only roundabout 5-6 \% of the transferred
characters are made for human consumption.{[}2{]}

Having a 95 \% overhead is rather undesirable for both the consumers and
the creators of the website. Since it is a widespread problem without a
definable point of failure one can argue `the platform itself' is the
failure. By design, every pageload results in a monolithic DOM tree
managed by the browser engine. Whether rendering just a bunch of static
text nodes or an ever changing webapp, the underlying global nature of
the DOM tree remains the same. Every additional piece of code added to
the webpage will invisibly add another fold of complexity to this global
object.

In a non-deterministic runtime environment, encapsulation and
modularization is a typical pattern to make complexity manageable and to
accommodate future uncertainty.{[}3, p. 1{]} Since years the average JS
payload is steadily rising which can be interpreted as a trend towards
more dynamic websites. The demands to the browser platform have changed
from a static page renderer to a \textbf{dynamic UI machine} without
significantly changing the underlying architecture. Under the current
situation only additional layers of abstraction can handle complexity.

In the recent years many \textbf{frameworks}, libraries and
methodologies approached the global nature of the DOM by scoping assets
and design rules into maintainable components. While the DOM cannot be
scoped, JS can. Many frameworks like React, Angular or Vue, just to name
a few, ditched the old rule of separated HTML, CSS and JS in favor of an
additional layer of abstracted JS components. Quite often those
frameworks mimic a MVC pattern on top of the browser engine which is a
reasonably simple pattern to build UIs. While frameworks are a valid
approach for building scalable web applications they remain highly
opinionated, embody inherent complexity themselves and can change and
break over time. Another drawback is code inflation which is a crucial
point for performance. All of those bottlenecks in the web demand for a
new standardized ways for creating and evolving complex web services.

In the year 2013 thinkers, creators and browser vendors joined together
to propose \emph{The Extensible Web Manifesto}.\footnote{\href{https://extensiblewebmanifesto.org/}{The
  Extensible Web Manifesto}} The claim of the manifesto was to enhance
the current web platforms with new low-level capabilities. Those
capabilities aimed to empower creators of the web to write more
declarative code and to abandon artificial abstractions. Four years
later, the enhancement of JavaScript leapfrogged and many new low-level
APIs were brought to life. With this new APIs at hand, a progressive web
developer can create robust websites with less code and less additional
libraries. This paper is an approach to unfold these
\textbf{browsernative} technologies to create overall simple and
resilient \textbf{microservices} for the browser

@@

In search of a better, simpler web architecture one might look on
already established patters proofed to fulfill enterprise needs.
Microservices are a good approach for tearing big monolithic systems
into fine-grained simple services with explicit defined boundaries. In a
nutshell a microservice is a small, autonomous service that works
together with other services seamlessly.{[}4, p. 2{]} Or with the words
of Fowler and Lewis: ``\ldots{} the microservice architectural style is
an approach to developing a single application as a suite of small
services, each running in its own process and communicating with
lightweight mechanisms, often an HTTP resource API.''{[}5{]} Yet at this
point a reader might spot some similarities with microservices and the
browser-based development: Both wrestle with the problem of monolithic
architecture and both use lightweight communication mechanisms. In fact,
many big companies of `the web' like Amazon or Netflix successfully
transformed their monolithic system into a service based system which
gives a taste of the power behind microservices.{[}5{]}

Microservices incorporate a wide array of ideas from developing scalable
software like \textbf{domain-driven design} to pursuing the real world
structure in the code.{[}4, p. 2{]} Another concept focuses on
\textbf{continuous delivery} for pushing software rapidly through
\textbf{automated deployment} mechanisms into production.{[}5{]}
Furthermore, microservices transcend the technical perspective and reach
into team organization.

As a primary source of comprehensive information this paper relies on
the work of Sam Newman{[}4{]} and the work of Fowler and Lewis{[}5{]}.
The purpose of this section is to develop confidence about the
architecture of microservices in the context of the browser platform.

\subsection{2.1. Componentization via
services}\label{componentization-via-services}

``A~\textbf{component}~is a unit of software that is independently
replaceable and upgradeable.'' {[}5{]} Components are the building
blocks of microservices. And microservices are the building blocks of
applications. Essentially the difference between microservices and
components is just the level of abstraction. Whether a concrete
microservice or a much more generic component, both share a similar set
of principles. Therefore this paper refers to both parts when talking
about \textbf{services}.

The first principle of services is the \textbf{loose coupling
principle}: changing and deploying one service should not result in
changing other parts of the system.{[}4, p. 30{]} Picking CSS for
example, where variables tend to be shadowed by higher specific values
making it increasingly difficult to keep changes, ought to merely affect
one place in the application. Therefore a \emph{browsernative
microservice} is expected to push encapsulation and to avoid variable
mutations outside its scope as much as possible. One solution would be
scoped CSS / JS to avoid variables leaking into the global namespace, as
well as to hide implementation details in order to avoid mutating
values.

The second principle of services is the \textbf{high cohesion
principle}: Whether designing a microservice or its components, we want
related behavior sit together and unrelated behavior to sit
elsewhere.{[}4, p. 30{]} High cohesion can be enhanced towards the more
dynamic \textbf{Single Responsibility Principle}: ``Gather together
those things that change for the same reason and separate those things
that change for different reasons.''{[}6{]} In a very quick and dirty
code quality analysis, the quality can be measured just by counting the
code changes which occur in order to implement a new functionality. An
arbitrary threefold MVC system should require a maximum of three changes
to implement or change functionalities. The problem in browser based
development is not only the global paradigm which makes changes
deliberately unpredictable. The high cohesion principle is violated by
the traditional separation along the siloed entities HTML, JS and CSS.
Understanding the relation of HTML markup towards another CSS file
generates incidental complexity. Different approaches emerged over the
recent years to join these entities. A \emph{browsernative service}
sought for a combination of the web native trinity HTML, JS and CSS.

Traditional web development relies on libraries to enhance the service
capabilities of the web platform. Compared to libraries a component
service offers multiple advantages for building, deployment and
shipping. As expressed earlier \emph{components for the web}, or web
components, are self-contained which means they embody all needed
functionality to get their job done. Therefore they have a much better
evolution mechanism in the service contracts. Changing functionality
will not break other services. A component can be progressively enhanced
which guarantees functionality throughout different versions whereas a
library is only loosely coupled to the implementation and therefore hard
to track in functionality. Changing a library may result in an
unforeseen amount of time fixing implementations. It is not unusual to
see websites embodying different versions of the same library to
guarantee functionality which lowers page performance significantly.

Another issue where web components stand out is related to performance
and especially the critical first page rendering. Libraries for the
browsers are traditionally `shipped' as non static immediately-invoked
functions. Following the Google RAIL model, a user-centric performance
measurement, a page load ought to take less than 1 second to catch up
users' attention.{[}7{]} There are many ways to optimize the critical
first render but as a rule of thumb a build-in web component might be
always superior to libraries in terms of first rendering. The
fine-grained lifecycle methods which will be described in the technical
section of this paper give the developer far reaching optimization
opportunities.

The last argument in favor of components over libraries is the more
explicit interface.{[}5{]} While the functionality of a library needs
documentation to be accessible, a component functionality is exposed via
the components' signature. HTML markup is an expressive syntax and
therefore convenient for steering a web component using merely
attributes and values.

\subsection{2.2. Organized around business
capabilities}\label{organized-around-business-capabilities}

\begin{quote}
``organizations which design systems \ldots{} are constrained to produce
designs which are copies of the communication~structures~of these
organizations''. {[}8{]}
\end{quote}

Emphasizing the human factor in microservices is a key feature.
Microservices are a product of real-world usage.{[}4, p. 1{]} Instead of
splitting the team along the technology stack (UI Experts
-\textgreater{} Middleware -\textgreater{} Database) a microservice
approach models teams around \textbf{business capabilities}.{[}5{]}
Consequently every team is capable of planning, designing, implementing,
testing and maintaining their very own microservice. Along the
technology stack every member gains high competence in the service
architecture which can be positive for evolving the service over its
whole lifecycle.

Real-world domains tend to be complex and multifaceted. To unfold their
complexity, domains can be subdivided into \textbf{bounded
context}.{[}4, p. 31{]} For example, customer service is a business
domain but with varying bounded contexts. One context can be sales,
another context could be support etc. Every context makes different
assumptions about the underlying model and draws an explicit interface
where it decides what to share with other contexts.{[}4, p. 30{]}
Evaluating each bounded context within each business domain will
eventually shape the data persistency model likewise the interface to
the service. This methodology can be iterated over and over again. A
sales service for example might be evaluated to different sales contexts
resulting in differing interfaces on differing devices shaped by
browsernative microservices. By dividing the service into clear defined
business boundaries it becomes easier to define a smart API of the
service.

Assigning service responsibility to a team, the so called
\textbf{Definition of Done} shifts from `accomplishing projects' to
`accomplishing products'. This new paradigm not only changes the
administrative overhead like budgeting or resource allocation.
Furthermore, it creates a kind of responsibility connection from the
team to the service which can be best described as \textbf{shared
governance} model. ``Each team collectively share responsibility for
evolving the technical vision of the system.''{[}4, p. 247{]}
Expectedly, those teams are more motivated within their very own service
and exhibit a more sophisticated iteration time.{[}5{]}

For many companies working in the spheres of the internet, the client
side is highly important for their business. In fact, business goals and
capabilities can be derived from frontend needs. The state of the web is
not only a story of numerous artifacts, it is also a story of a highly
fragmented market along devices, operating systems, differing sizes and
functionalities. Different devices again have different assumptions
about the technology stack. Splitting teams along the stack results in
an slow paced back and forth negotiation about every change to be made
can result in prolonged and therefor expensive iteration. As browser
technologies, design guidelines and devices change rapidly it makes
sense to shift responsibility towards the team altogether.

@@

To ensure the microservice functionality among teams and different
services requires thoughtful decentralization. Emphasizing once more the
real-world capabilities of microservices a message channel architecture
can be derived from patterns known from traditional postal services. A
physical letter has only two smart endpoints, entitled to read and
process the message while packaging as well as the connection itself is
mostly standardized. \textbf{Smart endpoints and dumb pipes} is coined
to the approach of designing communication mostly decoupled and as
cohesive as possible.{[}5{]} Applying these rules to the web platform
can result into building unified JSON message objects passed along
`dumb' middleware components. A typical browser event comes close to
this definition and suits arguably well for in-memory communication
between web components and microservices.

Microservices heavily rely on simple HTTP request-response with resource
APIs and lightweight messaging.{[}5{]} Newman recommends
technologic-agnostic REST APIs to free data persistence from
implementation constraints.{[}4, p. 247{]} The advantage of this overall
simple communication model is the suitability for both frontend-backend
likewise backend-backend communication. A service therefore can evolve
from a heavy backend with a lot of network roundtrips to a leaner
backend seamlessly. The browser build-in \textbf{fetch API} which is
essentially a HTTP request can be heavily incorporated into a
browsernative microservice to ensure communication to services in the
backend.

\subsection{2.4. Decentralized
governance}\label{decentralized-governance}

Microservices are separate entities and decentralization is important to
ensure autonomy. This paper already described fragmented services
bounded to singular business context, choreographed by simple
communication protocols developed and evolved by autonomous teams.

This distributed nature empowers teams to create their own technology
stack, tools and services designed in the spirit of language- and
platform independence and to share their knowledge across other
parties.{[}5{]} In the recent years many big companies like Facebook,
Google, Netflix and others followed that spirit and published their
ideas and implementations open source. The previously mentioned React
for example is a product of Facebook's need to ensure a consistent
frontend experience. In fact, many tools and techniques are byproduct of
vital interaction of specific domain problems and their solutions.

The spirit of freedom cannot be applied universally to
\emph{browsernative microservices} as the browser and its underlying DOM
will, to a certain degree, be the limiting factor. Talking about the
browser, a reader might be falsely tempted to narrow his perception
towards the obvious VIEW layer only. In recent years the major browser
engines have grown to fully-fledged app deployment platforms offering
connectors to build-in databases, multithreading support and
ever-growing JS build-ins like speech synthesis or push notifications.
So-called \textbf{Progressive Web Apps}\footnote{\href{https://developers.google.com/web/progressive-web-apps}{Progressive
  Web Apps}}, a bunch of criteria for building good browser apps, can
achieve a similar look and feel like native apps. And last but not least
services like NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}
  \#\# 2.5. Decentralized data management} effectively compiling `the
web' to native machine code lower the boundary between native and
browser code even further.

JS is the widely accepted language of the web. Nevertheless, a
microservice engineering team might choose another language for various
reasons. Transpiling languages to JS as target language is a stable
solution nowadays. Languages like TypeScript, ClojureScript or
PureScript compile to JS even exclusively. Once web components hit a
critical mass there will be most likely some library support or foreign
function interface towards ES6 modules, which are mandatory in the new
specifications.

Another more real life decentralization aspect derives from the easiness
of deployment in a safe, sandboxed environment. Web components virtually
ship no overhead or require dedicated build tools. This makes them ideal
candidates for sharing and open source publishing similar to the largest
JS package registry NPM. In the spirit of NPM web components can be
perceived as frontend packages with an HTML interface instead of a JS
signature. Webcomponents.org is a registry for ready-to-use components
of every scale and purpose where even Google shares a lot of their
material design elements.

Data Management in a microservice follows the same modular philosophy as
the service implementation. As mentioned before, different bounded
contexts make different assumptions of the underlying models.
Decentralized decisions about conceptual models demand decentralized
data storage decisions.{[}5{]} Todays web architectures aim to leverage
an increasing amout of \textbf{processing to the client} to avoid
time-consuming roundtrips especially in mobile networks.{[}9{]} Since
network roundtrips are costly it is a good advice to only query as much
data as needed and cache as much as possible. The build-in LocalStorage
or its successor IndexedDB are mature persistence technologies and
libraries like PouchDB\footnote{\href{https://pouchdb.com/}{PouchDB}}
even offer adapters for syncing to the server out of the box.

``Microservices prefer letting each service manage its own
database.''{[}Fowler2014{]} Ben Issa, chief architect of ING Australia
emphasizes this pragmatism on APIs in a conference talk. At ING the
frontend demands tailor the backend APIs. APIs may be produced
automatically and not even Issa knows how many APIs exists.{[}10{]} At
ING, they are using a pattern called \textbf{backends for frontends}
empowering the team to craft their UI and backend in a one-to-one
relationship.{[}4, p. 72{]}

To see this pattern in the field a reader might have a look at
Facebook's GraphQL\footnote{\href{http://graphql.org/}{GraphQL}}.
GraphQL is a query language for the frontend. The backend solely replies
to the frontend needs. Another well documented example in the field is
Cognitects Datomic\footnote{\href{http://www.datomic.com/}{Datomic}},
where parts of the database will be reflected to the client. A so-called
Transactor ensures ACID compliance.
